---
title: "R5 Rec-Shed Dashboard - Prototype"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme:
      primary: '#217a38'
      secondary: '#24583b'
      success: '#96d4ac'
      info: '#2f5b75'
      warning: '#ffc734'
      danger: '#892f31'
      base_font:
        google: Nunito
      code_font:
        google: Fira Code
      heading_font:
        google: Nunito
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(sf)
library(here)
library(dplyr)
library(tidyr)
library(leaflet)
library(ggplot2)
library(scales)
library(tigris)
library(RColorBrewer)
library(zipcodeR)
library(shiny)
library(janitor)
library(tidycensus)
library(stringr)
library(plotly)
# bslib::bs_themer()
```

```{r read in data}
# NOTEHD: saved data as rds files which are smaller

# spatial 
r5_au_bounds <- readRDS("r5_au_bounds.rds")
# NOTEHD: run to get data in global environment
# r5_au_bounds <- readRDS("dashboard/r5_au_bounds.rds")

# ridb
usfs_ridb2020and2021 <- readRDS("usfs_ridb2020and2021.rds")
# NOTEHD: run to get data in global environment
# usfs_ridb2020and2021 <- readRDS("dashboard/usfs_ridb2020and2021.rds")

us_geoms <- readRDS("df_state_geometries_us.rds")

# subset only 2020 and 2021 year
usfs_ridb2020and2021 <- usfs_ridb2020and2021 %>% 
  filter(year %in% c(2020, 2021))

# acs race
acs_race <- readRDS("race_acs_2017_2021.rds")
ca_acs_race <- readRDS("ca_acs_race.rds")
# acs_race <- readRDS("dashboard/race_acs_2017_2021.rds")

# rsconnect::configureApp("prototype-r5-dashboard", size="xlarge")

# zip code w population
zip_code_pop <- acs_race %>% 
  select(zip_code, zip_code_population)
```

## Sidebar {.sidebar}

```{r}
shiny::radioButtons("year", label = "Select a year:",
                          choices = sort(unique(usfs_ridb2020and2021$year)),
                          selected = 2021,
                          inline = TRUE)
```

```{r}
shiny::selectInput("forest", label = "Select a forest:",
                   choices = sort(unique(usfs_ridb2020and2021$forestname)),
                   selected = "Inyo National Forest")
```

**Metropolitan Group Demo Dashboard for USFS R5 RLT - Not for Public Use**

Select a year and a forest to change the visualizations on the right.

Dashboard prototype was developed by [Halina Do-Linh](https://github.com/hdolinh) and is based on the [Outdoor Equity App](https://github.com/outdoor-equity), a Shiny web app that explore patterns of visitor use of reservable overnight sites (such as campgrounds, cabins, hike-in, and more) to support equitable access to the outdoors.

## Row 1 {data-width=500} {data-height=400}

### R5 Map {.no-title}

```{r}
# mapview::mapview(r5_au_bounds)

# create subset of park info
park_sub_sf <- usfs_ridb2020and2021 %>% 
  dplyr::group_by(year,
                  forestname,
                  park,
                  sitetype,
                  facilityzip,
                  facilitylongitude,
                  facilitylatitude) %>%
  dplyr::summarize(n = n()) %>%
  # filter(is.na(sitetype) == TRUE) NOTEHD: Need to decide site types for NA sitetypes (e.g. xmas tree permits)
  dplyr::select(!n) %>%
  # create as sf object with 4326 crs
  st_as_sf(
    coords = c("facilitylongitude", "facilitylatitude"),
    crs = 4326,
    remove = FALSE
  )

# transform to 4326 crs
r5_bounds <- r5_au_bounds %>% sf::st_as_sf(crs = 4326)

reactive_r5_bounds <- shiny::reactive({
  r5_bounds %>%
    filter(forestname == input$forest)
})

reactive_park <- shiny::reactive({
  park_sub_sf %>% 
    filter(year == input$year) %>% 
    filter(forestname == input$forest)
})

leaflet::renderLeaflet({
  
  camping_icon <- makeIcon(
    "camping_location_icon.png",
    # here("images/camping_location_icon.png"), NOTEHD: run locally?
    iconWidth = 25,
    iconHeight = 25
  )
  
  leaflet() %>%
    addPolygons(
      data = reactive_r5_bounds(),
      weight = 1,
      popup = paste0(reactive_r5_bounds()$forestname)
    ) %>%
    addMarkers(
      lng = reactive_park()$facilitylongitude,
      lat = reactive_park()$facilitylatitude,
      popup = paste0(reactive_park()$park,
                     "<br>",
                     "Type of site: ",
                     reactive_park()$sitetype),
      icon = camping_icon
    ) %>% 
    addProviderTiles(providers$CartoDB.Voyager)
})
```

### About this Forest {data-width=250}

Add information about a specific forests and the reservable sites within it.


## Row 2 {data-height=600}

### Visits by State Map {.no-title}

```{r}

reactive_states_visits_map <- shiny::reactive({
  
# summarize ridb data
summary_state_vists <- usfs_ridb2020and2021 %>%
  filter(forestname == input$forest) %>% 
  group_by(year,
           forestname,
           customerstate) %>%
  summarize(count = n())

# join data
us_geoms %>% 
  left_join(summary_state_vists,
            by = c("state_abbrev" = "customerstate")) %>%
  st_transform('+proj=longlat +datum=WGS84') %>% 
  drop_na()
})

leaflet::renderLeaflet({
  # map
  bins <- c(0, 10, 50, 230, 500, 800, 1000, 5000, 140000)
  pal <- colorBin("BrBG", domain = reactive_states_visits_map()$count, 
                  bins = bins)
  
  leaflet() %>%
    setView(lat = 49.850033,
            lng = -125.6500523,
            zoom = 3) %>%
    addPolygons(
      data = reactive_states_visits_map(),
      weight = 1,
      opacity = 2,
      fillColor = ~ pal(count),
      fillOpacity = 0.7,
      color = "white",
      popup = paste0(
        reactive_states_visits_map()$state_abbrev,
        "<br>",
        "Total number of visits: ",
        scales::number_format(big.mark = ",")(reactive_states_visits_map()$count)
      )
    ) %>%
    addLegend(
      "topright",
      title = "Number of Visits",
      pal = pal,
      values = bins,
      opacity = 1
    ) %>%
    addProviderTiles(providers$CartoDB.Voyager)
})
```

### Visits by State {data-width=250}

This map displays the number of reservations made to the selected National Forest by state. Click on a state to display a pop-up that shows the exact number of reservations or visits made from that state to the selected National Forest.

## Row 3 {data-height=600}

### Visits by CA ZIP Codes {.no-title}

```{r}
# ca_zips <- tigris::zctas()
# 
# # 1,763 zips in CA according to ACS
# zip_ca <- zipcodeR::zip_code_db %>% 
#   filter(state == "CA") %>% 
#   filter(zipcode_type == "Standard") %>% 
#   select(zipcode,
#          major_city,
#          county)
# 
# joined_zips <- left_join(zip_ca, ca_zips, by = c("zipcode" = "ZCTA5CE20")) 
# 
# # 38 zips NA meaning no geoms
# # test_zips <- joined_zips %>% 
# #   filter(is.na(GEOID20) == TRUE)
# 
# clean_ca_zips <- joined_zips %>% 
#   filter(is.na(GEOID20) == FALSE) %>% 
#   select(zipcode,
#          major_city,
#          county,
#          geometry) %>% 
#   st_as_sf() %>% 
#   st_transform(4326) %>% 
#   rmapshaper::ms_simplify(keep = 0.005, keep_shapes = TRUE)
# 
# # mapview::mapview(clean_ca_zips)
# 
# ridb_zip_visits <- usfs_ridb2021 %>% 
#   select(year,
#          forestname,
#          customerzip)
# 
# ca_zip_visits_map <- left_join(clean_ca_zips, ridb_zip_visits,
#                                by = c("zipcode" = "customerzip"))
# # count num of visits by zip code
# ca_zip_visits_map <- ca_zip_visits_map %>% 
#   group_by(year,
#            forestname,
#            zipcode,
#            major_city,
#            county) %>% 
#   summarize(count = n())
# 
# saveRDS(ca_zip_visits_map, "dashboard/ca_zip_visits_map.rds")

ca_zip_visits_map <- readRDS("ca_zip_visits_map.rds")
# NOTEHD: run in rmd
# ca_zip_visits_map <- readRDS("dashboard/ca_zip_visits_map.rds")

# reactive df
reactive_ca_zip_visits_map <- shiny::reactive({
  
  ca_zip_visits_map %>%
    sf::st_transform('+proj=longlat +datum=WGS84') %>% 
    left_join(zip_code_pop, by = c("zipcode" = "zip_code")) %>% 
    filter(year == input$year) %>% 
    filter(forestname == input$forest)
})

reactive_forest_ref <- shiny::reactive({
  
  r5_au_bounds %>%
    filter(forestname == input$forest) %>%
    st_transform('+proj=longlat +datum=WGS84')
  
})



leaflet::renderLeaflet({
  
  bins <- c(0, 5, 10, 30, 55, 100, 800, 1000, 3000)
  pal <- colorBin("BrBG", domain = reactive_ca_zip_visits_map()$count, 
                  bins = bins)

leaflet() %>% 
  setView(lat = 37.5000,
          lng = -119.4179,
          zoom = 6) %>% 
  addPolygons(data = reactive_forest_ref(),
              weight = 1,
              opacity = 1,
              color = "red") %>% 
  addPolygons(data = reactive_ca_zip_visits_map(),
              weight = 0.5,
              opacity = 1,
              fillColor = ~ pal(count),
              fillOpacity = 0.8,
              color = "black",
              popup = paste0(
                "<strong>", "Visitor Home ZIP: ",
                reactive_ca_zip_visits_map()$zipcode, "</strong>",
                "<br>",
                reactive_ca_zip_visits_map()$major_city,
                "<br>",
                reactive_ca_zip_visits_map()$county,
                "<br>",
                "Total number of visits: ",
                scales::number_format(big.mark = ",")(reactive_ca_zip_visits_map()$count)
              )
              ) %>% 
  addLegend(
    "topright",
    title = "Number of Visits",
    pal = pal,
    values = bins,
    opacity = 1
  ) %>% 
  addProviderTiles(providers$CartoDB.Voyager)
  
})

```

### Visits by CA ZIP Codes {data-width=250}

National Forests in Region 5 are primarily visited by California residents, this is no surprise, but it pushes us to further explore the RIDB data to see where in California visitors are traveling from. To do this, we combine RIDB data with ACS Census data.

This map shows the number of visits made to the selected National Forest by ZIP code. Click on a ZIP code boundary to display a pop-up that lists the Visitor's Home ZIP code, city, county, and the total number of visits made from the ZIP code.

The geometry in red is the selected National Forest and is added to the map as a reference point.

## Row 4 {data-width=300}{data-height=400}

### Estimated Racial Percentages of California Residents vs. Visitors

```{r}
# reactive
reactive_race <- reactive({
  
  ca_race_calc <- ca_acs_race %>%
    mutate(
      percent = (estimate / total_population_est),
      lower_bound = estimate - moe,
      upper_bound = estimate + moe,
      lower_percentage = (lower_bound / total_population_est),
      upper_percentage = (upper_bound / total_population_est),
    )
  
  ca_race_calc_pct <- ca_race_calc %>%
    select(zipcode,
           race,
           total_population_est,
           percent) %>%
    pivot_wider(names_from = "race",
                values_from = "percent")
  
  # join with ridb
  ridb_race <- left_join(usfs_ridb2020and2021,
                         ca_race_calc_pct,
                         by = c("customerzip" = "zipcode"))
  
  # CA population
  race_ca <- ca_race_calc_pct %>%
    summarize(
      white = weighted.mean(white, total_population_est,
                            na.rm = TRUE),
      black = weighted.mean(black,  total_population_est,
                            na.rm = TRUE),
      asian = weighted.mean(asian, total_population_est,
                            na.rm = TRUE),
      multiracial = weighted.mean(multiracial, total_population_est,
                                  na.rm = TRUE),
      other = weighted.mean(other, total_population_est,
                            na.rm = TRUE),
      native_american = weighted.mean(native_american, total_population_est,
                                      na.rm = TRUE),
      pacific_islander = weighted.mean(pacific_islander,
                                       total_population_est,
                                       na.rm = TRUE),
      hispanic_latinx = weighted.mean(hispanic_latinx, total_population_est,
                                      na.rm = TRUE)
    ) %>%
    pivot_longer(cols = 1:8,
                 names_to = "race",
                 values_to = "race_percent_average") %>%
    mutate(race = str_replace(
      string = race,
      pattern = "_",
      replacement = " "
    ),
    race = str_to_title(race))
  
  ridb_race <- ridb_race %>%
    filter(forestname == input$forest) %>%
    filter(year == input$year) %>%
    summarize(
      white = mean(white, na.rm = TRUE),
      black = mean(black, na.rm = TRUE),
      asian = mean(asian, na.rm = TRUE),
      multiracial = mean(multiracial, na.rm = TRUE),
      other = mean(other, na.rm = TRUE),
      native_american = mean(native_american, na.rm = TRUE),
      pacific_islander = mean(pacific_islander, na.rm = TRUE),
      hispanic_latinx = mean(hispanic_latinx, na.rm = TRUE)
    ) %>%
    pivot_longer(cols = 1:8,
                 names_to = "race",
                 values_to = "race_percent_average") %>%
    mutate(race = str_replace(
      string = race,
      pattern = "_",
      replacement = " "
    ),
    race = str_to_title(race))
  
  # join data for plotting
  left_join(
    x = ridb_race,
    y = race_ca,
    by = c("race"),
    suffix = c("_ridb", "_ca")
  ) %>%
    rename(RIDB = race_percent_average_ridb,
           CA = race_percent_average_ca) %>%
    pivot_longer(cols = 2:3,
                 names_to = "data_source",
                 values_to = "race_percent_average") %>%
    mutate(
      data_source = factor(data_source, levels = c("RIDB", "CA")),
      tooltip_start = case_when(
        data_source == "RIDB" ~ "Visitors live in communities with an estimated ",
        data_source == "CA" ~ ""
      ),
      tooltip_middle = case_when(
        data_source == "RIDB" ~ " ",
        data_source == "CA" ~ " of Californians are "
      ),
      tooltip_end = case_when(data_source == "RIDB" ~ " population",
                              data_source == "CA" ~ ""),
      race = str_replace(
        string = race,
        pattern = "Other",
        replacement = "Other Race(s)"
      ),
      race = factor(
        race,
        ordered = TRUE,
        levels = c(
          "Hispanic Latinx",
          "White",
          "Asian",
          "Black",
          "Multiracial",
          "Native American",
          "Pacific Islander",
          "Other Race(s)"
        )
      )
    )
  
}) # EO RDF

# parameters
groups_colors_ridb_ca <- c("RIDB" = "#64863C", "CA" = "#e8e8e8")
text_colors_ridb_ca <- c("RIDB" = "#466C04", "CA" = "#a3a3a3")

plotly::renderPlotly({
  
  x_max <-
  max(reactive_race()$race_percent_average) + 0.1 # max x rounded to nearest 5
  
  # plot for shiny app
  race_plotly <- ggplot(data = reactive_race()) +
    geom_col(aes(
      x = race_percent_average,
      y = reorder(race, desc(race)),
      fill = data_source,
      text = paste0(
        tooltip_start,
        scales::percent(race_percent_average, accuracy = 0.1),
        tooltip_middle,
        race,
        tooltip_end
      )
    ),
    position = "dodge") +
    scale_x_continuous(labels = percent, limits = c(0, x_max)) +
    scale_fill_manual(values = groups_colors_ridb_ca) +
    # geom_text(
    #   aes(
    #     x = race_percent_average,
    #     y = reorder(race, desc(race)),
    #     label = scales::percent(race_percent_average, accuracy = 0.1),
    #     col = data_source
    #   ),
    #   position = position_dodge(width = 1),
    #   size = 3
    # ) +
    scale_color_manual(values = text_colors_ridb_ca) +
    labs(x = "Percentage (%)",
         y = "") +
    theme_minimal()
  race_plotly
  
  ggplotly(race_plotly,
           tooltip = list("text")) %>%
    # style(textposition = "right") %>%
    layout(# title = list(text = paste0("Inyo National Forest", '<br>',
      #                                 'Estimated Racial Percentages of California Residents vs. Visitors'),
      #                   font = list(size = 15)),
      showlegend = FALSE) %>%
    config(
      modeBarButtonsToRemove = list(
        "pan",
        "select",
        "lasso2d",
        "autoScale2d",
        "hoverClosestCartesian",
        "hoverCompareCartesian"
      )
    )
  
})
```


### Estimated Racial Percentages of California Residents vs. Visitors {data-width=250}

This interactive plot compares the racial breakdown of California to the estimated racial breakdown of the selected National Forest. This can provide insight into whether or not a National Forest is attracting visitors that aligns with the racial diversity of California.

The **grey bars** represent the **California** racial breakdown, and the **green bars** represent the estimated racial breakdown of the selected **National Forest**. **Hover** over a bar to see **exact estimates**.

## Row 5 {data-width=300}{data-height=400}

### Distance Traveled by Race

```{r}
ca_race_calc <- ca_acs_race %>%
  mutate(
    percent = (estimate / total_population_est),
    lower_bound = estimate - moe,
    upper_bound = estimate + moe,
    lower_percentage = (lower_bound / total_population_est),
    upper_percentage = (upper_bound / total_population_est),
  )

ca_race_calc_pct <- ca_race_calc %>%
  select(zipcode,
         race,
         total_population_est,
         percent) %>%
  pivot_wider(names_from = "race",
              values_from = "percent")

# cal top race quartile
acs_race_top <- ca_race_calc_pct %>%
  pivot_longer(cols = 3:10,
               names_to = "race",
               values_to = "race_percentage") %>%
  drop_na(race_percentage)

# weighted median value (weighted based on ZIP code populations)
weighted_half <- weighted.mean(x = acs_race_top$race_percentage,
                               w = acs_race_top$total_population_est)

# drop rows below weighted median
df_half <- acs_race_top %>% filter(race_percentage >= weighted_half)

# weighted 3rd quartile -- weighted median value of top half (weighted based on ZIP code populations)
weighted_quartile <-  weighted.mean(x = df_half$race_percentage,
                                    w = df_half$total_population_est)

# join with ridb
ridb_race <- left_join(usfs_ridb2020and2021,
                       ca_race_calc_pct,
                       by = c("customerzip" = "zipcode"))

# reactive
plot_data <- shiny::reactive({
  
  ridb_race %>%
    # filter
    filter(forestname == input$forest) %>%
    filter(year == input$year) %>%
    # select to variables of interest
    select(
      park,
      customerzip,
      asian,
      black,
      hispanic_latinx,
      multiracial,
      native_american,
      other,
      pacific_islander,
      white,
      dist_travel_mi
    ) %>%
    pivot_longer(cols = 3:10,
                 names_to = "race",
                 values_to = "race_percentage") %>%
    # filter rows that fall above 3rd quartile value
    filter(race_percentage >= weighted_quartile) %>%
    # updated racial category name strings for plotting
    mutate(
      race = str_replace(
        string = race,
        pattern = "_",
        replacement = " "
      ),
      race = str_to_title(race),
      race = str_replace(
        string = race,
        pattern = "Other",
        replacement = "Other Race(s)"
      )
    ) %>%
    drop_na(dist_travel_mi, race_percentage) %>%
    # summarize to inner quartile range, median, and total reservations
    group_by(race) %>%
    summarize(
      median_distance_traveled_mi = median(dist_travel_mi),
      quartile_lower = quantile(dist_travel_mi)[[2]],
      quartile_upper = quantile(dist_travel_mi)[[4]],
      count = n()
    )
})

# plot
plotly::renderPlotly({
  plotly <- ggplot(data = plot_data(),
                   aes(
                     x = median_distance_traveled_mi,
                     y = reorder(race, median_distance_traveled_mi)
                   )) +
    geom_segment(aes(xend = 0, yend = race)) +
    geom_point(
      aes(
        text = paste0(
          scales::comma(count, accuracy = 1),
          " unique visits were made by people who live in ZIP codes<br>with high ",
          race,
          " populations. Typically these visitors traveled between<br>",
          scales::comma(quartile_lower, accuracy = 1),
          " and ",
          scales::comma(quartile_upper, accuracy = 1),
          " miles, with a median distance of ",
          scales::comma(median_distance_traveled_mi, accuracy = 1),
          " miles."
        )
      ),
      size = 3.5,
      shape = 21,
      stroke = 2,
      fill = "#64863C",
      color = "#64863C"
    ) +
    scale_y_discrete(expand = c(0.45, 0)) +
    labs(x = paste("Estimated Distance Traveled from Home to National Forest (miles)"),
         y = "") +
    theme_minimal() +
    theme(
      plot.background = element_rect("white"),
      panel.grid.major.y = element_blank(),
      legend.position = "none"
    )
  
  # create plotly
  ggplotly(plotly,
           tooltip = list("text")) %>%
    config(
      modeBarButtonsToRemove = list(
        "zoom",
        "pan",
        "select",
        "lasso2d",
        "autoScale2d",
        "hoverClosestCartesian",
        "hoverCompareCartesian"
      )
    ) %>%
    add_annotations(
      text = "Reservations from ZIP codes<br>with high proportions of:",
      x = -0.5,
      y = 0.9,
      font = list(size = 11),
      xref = 'paper',
      yref = 'paper',
      showarrow = FALSE
    )
})
```

### Distance Traveled by Race {data-width=250}

This interactive plot compares the distance traveled by race.

**Hover** over a point at the end of the bar to see **exact estimates** and more context for what that data point shows.

## Row 6 {data-width=300}{data-height=400}

### Number of Days in Advance Reservation is Made by Visitors with Different Racial Groups

```{r}
ca_race_calc <- ca_acs_race %>%
  mutate(
    percent = (estimate / total_population_est),
    lower_bound = estimate - moe,
    upper_bound = estimate + moe,
    lower_percentage = (lower_bound / total_population_est),
    upper_percentage = (upper_bound / total_population_est),
  )

ca_race_calc_pct <- ca_race_calc %>%
  select(zipcode,
         race,
         total_population_est,
         percent) %>%
  pivot_wider(names_from = "race",
              values_from = "percent")

# cal top race quartile
acs_race_top <- ca_race_calc_pct %>%
  pivot_longer(cols = 3:10,
               names_to = "race",
               values_to = "race_percentage") %>%
  drop_na(race_percentage)

# weighted median value (weighted based on ZIP code populations)
weighted_half <- weighted.mean(x = acs_race_top$race_percentage,
                               w = acs_race_top$total_population_est)

# drop rows below weighted median
df_half <- acs_race_top %>% filter(race_percentage >= weighted_half)

# weighted 3rd quartile -- weighted median value of top half (weighted based on ZIP code populations)
weighted_quartile <-  weighted.mean(x = df_half$race_percentage,
                                    w = df_half$total_population_est)

# join with ridb
ridb_race <- left_join(usfs_ridb2020and2021,
                       ca_race_calc_pct,
                       by = c("customerzip" = "zipcode"))

reactive_race_bw <- shiny::reactive({
  ridb_race %>%
    # filter
    filter(forestname == input$forest) %>%
    filter(year == input$year) %>%
    # select to variables of interest
    select(
      park,
      customerzip,
      asian,
      black,
      hispanic_latinx,
      multiracial,
      native_american,
      other,
      pacific_islander,
      white,
      bookingwindow
    ) %>%
    pivot_longer(cols = 3:10,
                 names_to = "race",
                 values_to = "race_percentage") %>%
    # filter rows that fall above 3rd quartile value
    filter(race_percentage >= weighted_quartile) %>%
    # updated racial category name strings for plotting
    mutate(
      race = str_replace(
        string = race,
        pattern = "_",
        replacement = " "
      ),
      race = str_to_title(race),
      race = str_replace(
        string = race,
        pattern = "Other",
        replacement = "Other Race(s)"
      )
    ) %>%
    drop_na(bookingwindow, race_percentage) %>%
    # summarize to inner quartile range, median, and total reservations
    group_by(race) %>%
    summarize(
      median_booking_window = median(bookingwindow),
      quartile_lower = quantile(bookingwindow)[[2]],
      quartile_upper = quantile(bookingwindow)[[4]],
      count = n()
    )
})

plotly::renderPlotly({
  plotly <- ggplot(data = reactive_race_bw(),
                   aes(x = median_booking_window,
                       y = reorder(race, median_booking_window))) +
    geom_segment(aes(xend = 0, yend = race)) +
    geom_point(
      aes(
        text = paste0(
          scales::comma(count, accuracy = 1),
          " unique visits were made by people who live in ZIP codes<br>with high ",
          race,
          " populations. Typically these visitors reserved their visit between<br>",
          scales::comma(quartile_lower, accuracy = 1),
          " and ",
          scales::comma(quartile_upper, accuracy = 1),
          " days before the start of their trip, with a median booking window of ",
          scales::comma(median_booking_window, accuracy = 1),
          " days."
        )
      ),
      size = 3.5,
      shape = 21,
      stroke = 2,
      fill = "#64863C",
      color = "#64863C"
    ) +
    scale_y_discrete(expand = c(0.45, 0)) +
    labs(
      x = paste("Estimated Number of Days in Advance Reservation is Made (days)"),
      y = ""
    ) +
    theme_minimal() +
    theme(
      plot.background = element_rect("white"),
      panel.grid.major.y = element_blank(),
      legend.position = "none"
    )
  
  # create plotly
  ggplotly(plotly,
           tooltip = list("text")) %>%
    config(
      modeBarButtonsToRemove = list(
        "zoom",
        "pan",
        "select",
        "lasso2d",
        "autoScale2d",
        "hoverClosestCartesian",
        "hoverCompareCartesian"
      )
    ) %>%
    add_annotations(
      text = "Reservations from ZIP codes<br>with high proportions of:",
      x = -0.15,
      y = 0.9,
      font = list(size = 11),
      xref = 'paper',
      yref = 'paper',
      showarrow = FALSE
    )
})
```

### Booking Window by Race {data-width=250}

This interactive plot compares the booking window by race.

**Hover** over a point at the end of the bar to see **exact estimates** and more context for what that data point shows.

## Row 7 {data-width=300}{data-height=400}

### Reservations to Different Site Types Made by Non-white Visitors

```{r}
ca_race_calc <- ca_acs_race %>%
  mutate(
    percent = (estimate / total_population_est),
    lower_bound = estimate - moe,
    upper_bound = estimate + moe,
    lower_percentage = (lower_bound / total_population_est),
    upper_percentage = (upper_bound / total_population_est),
  )

ca_race_calc_pct <- ca_race_calc %>%
  select(zipcode,
         race,
         total_population_est,
         percent) %>%
  pivot_wider(names_from = "race",
              values_from = "percent")

# cal top race quartile
acs_race_top <- ca_race_calc_pct %>%
  pivot_longer(cols = 3:10,
               names_to = "race",
               values_to = "race_percentage") %>%
  drop_na(race_percentage)

# weighted median value (weighted based on ZIP code populations)
weighted_half <- weighted.mean(x = acs_race_top$race_percentage,
                               w = acs_race_top$total_population_est)

# drop rows below weighted median
df_half <- acs_race_top %>% filter(race_percentage >= weighted_half)

# weighted 3rd quartile -- weighted median value of top half (weighted based on ZIP code populations)
weighted_quartile <-  weighted.mean(x = df_half$race_percentage,
                                    w = df_half$total_population_est)

# join with ridb
ridb_race <- left_join(usfs_ridb2020and2021,
                       ca_race_calc_pct,
                       by = c("customerzip" = "zipcode"))

reactive_race_sitetype <- shiny::reactive({
  ridb_race %>%
    # filter
    filter(forestname == input$forest) %>%
    filter(year == input$year) %>%
    # select to variables of interest
    select(
      park,
      customerzip,
      asian,
      black,
      hispanic_latinx,
      multiracial,
      native_american,
      other,
      pacific_islander,
      white,
      sitetype
    ) %>%
    pivot_longer(cols = 3:10,
                 names_to = "race",
                 values_to = "race_percentage") %>%
    # filter rows that fall above 3rd quartile value
    filter(race_percentage >= weighted_quartile) %>%
    # updated racial category name strings for plotting
    mutate(
      race = str_replace(
        string = race,
        pattern = "_",
        replacement = " "
      ),
      race = str_to_title(race),
      race = str_replace(
        string = race,
        pattern = "Other",
        replacement = "Other Race(s)"
      )
    ) %>%
    filter(race != "White") %>%
    drop_na(sitetype, race_percentage) %>%
    # summarize to total reservations for each site type
    count(race, sitetype) %>%
    rename(count = n) %>%
    # filter to indicated site type and update string for plotting
    mutate(
      sitetype = str_to_title(sitetype),
      sitetype = str_replace(
        string = sitetype,
        pattern = "Rv",
        replacement = "RV"
      )
    )
})

plotly::renderPlotly({
  plotly <- ggplot(data = reactive_race_sitetype()) +
    geom_col(aes(
      x = count,
      y = reorder(race, count),
      fill = reorder(sitetype, count),
      text = paste0(
        "Of visits to ",
        sitetype,
        " overnight reservable sites, ",
        scales::comma(count, accuracy = 1),
        " reservations were made by <br>people who live in ZIP codes with high ",
        race,
        " populations."
      )
    )) +
    scale_y_discrete(expand = c(0.45, 0)) +
    labs(x = paste("Number of Reservations"),
         y = NULL,
         fill = NULL) +
    theme_minimal() +
    theme(
      plot.background = element_rect("white"),
      panel.grid.major.y = element_blank()
    )
  
  # create plotly
  ggplotly(plotly,
           tooltip = list("text")) %>%
    config(
      modeBarButtonsToRemove = list(
        "zoom",
        "pan",
        "select",
        "lasso2d",
        "autoScale2d",
        "hoverClosestCartesian",
        "hoverCompareCartesian"
      )
    ) %>%
    add_annotations(
      text = "Reservations from ZIP codes<br>with high proportions of:",
      x = -0.15,
      y = 1,
      font = list(size = 11),
      xref = 'paper',
      yref = 'paper',
      showarrow = FALSE
    )
})
```

### Reservations to Different Site Types Made by Non-white Visitors {data-width=250}

This interactive plot compares site type by non-white races.

**Hover** over a section of the bar to see **exact estimates** and more context for what that data point shows. Double-click the legend to hide and show different site types.
